Designing Types for R, Empirically

### 2020-05-15 paper/main.tex
Add some text for first evaluation section

%
%
%
%
%
%
\section{Evaluation Results}
\label{sec:evaluation}

@@

\AT{Aviral}

>>>>

%
%
%
%
%
%
\section{Evaluation}
\label{sec:evaluation}

@@

\subsection{Is our simple type language expressive enough to type R code?} Functions in
R can support arguments of multiple types by reflecting on the type to perform
appropriate conversions and computations. On the one hand, this form of ad-hoc
polymorphism makes R functions expressive; on the other hand, it requires
support for polymorphism in the typing scheme, making the type system complex. 
Our system lacks explicit support for polymorphism. Instead, it supports union and any
types. A union type is inferred when values of multiple types are observed at a
function parameter and return position. any type is inferred when an argument is
missing or unused. Furthermore, list element types can be inferred as any if
they have more than 5 different types. Values with many different types can
result in a large number of alternatives in the type inferred for that position,
making it hard to read and unwieldy for end-users. To evaluate this
''expressivity gap`' between our type inference scheme and R functions,
we look at the number of different types in the parameter and return types of
the XXX functions in our corpus.

We observe that XXX% of the positions have a single type ascribed to them. XXX%
have five or more types ascribed to them. Parameter PARAMNAME of function
XXXXCOLONCOLONXXX has the maximum, XXX, types ascribed to it. This implies that
most functions admit or return a uni-typed value.

Figure XXX shows ten most common type categories that occur by themselves. We
observe that three scalar types, double, logical, and character are most
frequent. This is followed by XXX vararg parameter positions. Next, XX% of the
parameter types are any because arguments at those positions were either missing
or unused in at least one call to those functions. Together, these types amount
to more than 50%. This is followed by double vector, the unit type, null,
character vector, and classes matrix and data.frame respectively.

Figure 1 shows the number of alternatives in the inferred type of function
parameter and return values.

XXX% of the parameter positions are monomorphic. 
\typetracer observes XXX calls to XXX functions with XXX arguments. This
translates to XXX parameters and XXX return types. 
A parameter or return type is  from unioning all the distinct observed types of

Figure 1 shows the distribution of parameter and return types for these
functions.

class is the most common type in our corpus, accounting for XXX% of all the
inferred types. This is followed by an almost equal proportion,
28%, scalar types and, surprisingly, only 13% vector types. Vectors are the
workhorse of R, most primitive operations in R are vectorized and literals are
treated as single element vectors. It is unexpected to observe twice as many
scalars as vectors. It is even more unexpected to observe class as the most
common type. class of an object is used for dynamic dispatch by the S3 object
system in R. This suggests that usage of S3 is prevalent in our corpus.

null types. Together, these four types make up for 80% of the observed types in
our corpus.
The other types, occurring only XXX% of the times include XXX environment types,
XXXX expression types, XXXX pairlist types and lastly, XXX externalptr types.
